A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-03-07 paper/paper1/ms.tex
Some text on Orbital shear

\subsection{Critical examination of the Parker Wind model}
An even more existential limit arises when we consider the assumptions of the isothermal Parker wind model, and the extreme geometry of \hatpb.  Figure \ref{fig:murrayClaySchematic} shows a schematic modeled after Figure 8 from \citet{2009ApJ...693...23M}.  The point at which the gas velocity exceeds the sound speed, the so-called sonic point, resides at 1.2 $R_p$, a mere $\sim9$ atmospheric scale heights from the planet surface.  The smallness of this radius can be understood from the significant effect of tidal gravity and the .  The pressure level here would be in the fractions of a millibar level, with an atmospheric temperature $\le T_\mathrm{eq}\sim10^3$~K and not yet the $T\sim10^4$~K exosphere range.  Depending on the mass loss rate, the transition from primarily neutral to primarily ionized plasma may occur outside or inside the sonic radius.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/murray_clay_schematic_v0p1.png}
    \caption{Schematic of the relevant length scales in a stratified atmosphere, modeled after Figure 8 from \citet{2009ApJ...693...23M}.  The sonic point resides at a mere 2 atmospheric scale heights, and well below the photoionization level of Hydrogen, raising the possibility that the supersonic flow begins neutral and becomes ionized later on.}
    \label{fig:murrayClaySchematic}
\end{figure}

Eventually, for very large mass loss rates, the width of the line begins to grow.

the observations indicate a 3D geometry of the outflow-- a preference for escaping material to overtake the planet along the orbital path.  Once the The 1D Parker wind models for high mass loss rates
During the moment of the planet transit midcenter ($\phi=0.0$), the subsolar point on the planet dayside will have material be preferrentially removed from the line-of-sight

>>>>

\subsection{Causes for the leading tail}
Several physical phenomena could give rise to a leading tail geometry.  Here we explore two leading candidates: orbital shear and stellar wind confinement.

Figure \ref{fig:KeplerianShear} shows an illustration of Keplerian shear adapted to the system properties of \hatpb.  In this scenario, the wind launches primarily from the dayside, with relatively little or no wind launched from the nightside.  The wind initially launches radially from the planet, with the strongest wind located near the sub-stellar point, the line connecting the planet to star along the vertical axis in the figure.  Inefficient heating near the terminator subdues the mass loss in the $x-$direction, meaning that this wind exhibits not a hemi-spherical shape, more like a bulb or torpedo.  The gas increasingly experiences the star's Keplerian potential past the Roche lobe, accelerating in the direction of orbital motion, $+x$.  The accelerating column eventually overtakes the planet completely, with the prospect of extending to hundreds of planetary radii.

Stellar wind confinement...

\begin{figure}
    \includegraphics[width=0.8\linewidth]{figures/KeplerianShear_v0p3.png}
    \caption{Schematic of Keplerian orbital shear.  Locations outside the Roche lobe experience orbital shear from the Keplerian potential.  A wind launched primarily from the dayside will tend to form a leading tail.}
    \label{fig:KeplerianShear}
\end{figure}

\subsection{Critical examination of the Parker Wind model}

\begin{figure*}[t]
    \includegraphics[width=\linewidth]{figures/murray_clay_schematic_v0p2.png}
    \caption{Schematic of the relevant length scales in a stratified atmosphere, modeled after Figure 8 from \citet{2009ApJ...693...23M}.  The sonic point resides at a mere 2 atmospheric scale heights, and well below the photoionization level of Hydrogen, raising the possibility that the supersonic flow begins neutral and becomes ionized later on.}
    \label{fig:murrayClaySchematic}
\end{figure*}

An even more existential limit arises when we consider the assumptions of the isothermal Parker wind model, and the extreme geometry of \hatpb.  Figure \ref{fig:murrayClaySchematic} shows a schematic modeled after Figure 8 from \citet{2009ApJ...693...23M}.  The point at which the gas velocity exceeds the sound speed, the so-called sonic point, resides at 1.2 $R_p$, a mere $\sim9$ atmospheric scale heights from the planet surface.  The smallness of this radius can be understood from the significant effect of tidal gravity and the .  The pressure level here would be in the fractions of a millibar level, with an atmospheric temperature $\le T_\mathrm{eq}\sim10^3$~K and not yet the $T\sim10^4$~K exosphere range.  Depending on the mass loss rate, the transition from primarily neutral to primarily ionized plasma may occur outside or inside the sonic radius.
